---
title: "Zero-Truncated Negative Binomial Regression"
author: 'Group 6: Zi Wang, Dong Ding and Junfeng Luo'
date: "11/24/2018"
output: html_document
---
## Introduction to Zero-Truncated Negative Binomial Regression
Zero-truncated Negative Binomial Regression is used to model count data 
for which the value zero cannot occur and for which over dispersion exists. 
There are a lot of response variables that cannot have a value of 0, 
such as the duration patients are in hospital and the age of an animal. 
When people want to use regression on these count variables, 
they may want to use Negative Binomial Regression first 
because it is a useful model for the count data. 
However, it is the underlying assumption of Negative Binomial distributions 
that may cause a problem as these distributions allow zeros 
within their range of possible values. If the mean of the response is small, 
and it does not contain zeros,then the estimated parameters and standard errors
obtained by GLM may be biased, which means 
the Negative Binomial Regression model does not fit well. 
In this situation, the Zero-Truncated Negative Binomial Regression model 
can be used to solve this problem.

### Data Background
The data used in this tutorial is the Abalone Dataset, which comes from 
an original study about The Population Biology of Abalone by Warwick J Nash, 
Tracy L Sellers, Simon R Talbot, Andrew J Cawthorn and Wes B Ford (1994). 
The link of the data: https://www.kaggle.com/rodolfomendes/abalone-dataset . 
We will focusing on the following variables: 
Rings(Can give the age of the Abalone in years by adding 1.5), 
Sex(M, F, and I (infant)) and Length(Longest shell measurement). 
The response variable is Rings.

## R

Load the data into R using `read.csv`
```{r import, echo = TRUE,message = FALSE}
library(VGAM)
library(tidyverse)
library(MASS)
abalone_full = read.csv("abalone.csv")
```

Use the function `distinct` from `dplyr` package to select the variable 
that we want and remove replicate rows. 
Show the dimension and the summary of the new data named *abalone*. 
```{r select, echo = TRUE}
abalone = abalone_full %>%
  distinct(Rings, Sex, Length)
dim(abalone)
```

```{r summary_2, echo = TRUE}
summary(abalone)
```

Show the first 10 rows of the data *abalone*:
```{r show_1, echo = TRUE}
abalone[1:10,]
```

Now, we are going to visualize the response *Rings* to test whether our dataset 
is suitable to use the Zero-Truncated Negative Binomial Regression model.   
At first, from the summary of the data and the decription of the variables, 
the response *Rings* is a count variable that cannot have a value of 0.  
In addition, show the mean, standard deviation and histogram 
of the response *Rings*:
```{r show_2, echo = TRUE}
sprintf("The Mean of Rings = %4.3f, and the SD of Rings = %4.3f", 
        mean(abalone$Rings), sd(abalone$Rings))
```

```{r hist, echo = TRUE}
hist(abalone$Rings, xlab = "Rings", main = "Histogram of Rings")
```
  
The results show that the variable *Rings* has small means, 
and exists over dispersion.
Therefore, the Zero-Truncated Negative Binomial Regression model 
is suitable for modeling our dataset.


Use function `vglm` in `VGAM` package to use 
Zero-Truncated Negative Binomial Regression on the data *abalone* 
```{r TNB_1, echo = TRUE}
t_nb1= vglm(Rings ~ Sex + Length, 
            family = posnegbinomial(), data = abalone)
summary(t_nb1)
```
The Log-likelihood of the model is -4294.895

Fit the Negative Binomial Regression using `glm.nb` in `MASS` package 
on the data *abalone*
```{r NB_1, echo = TRUE}
nb1 = glm.nb(Rings ~ Sex + Length, data = abalone)
summary(nb1)
```
The Log-likelihood of the model is -8590.748/2 = -4295.874  

Compare these two models using Log-likelihood:  
Because -4294.895 > -4295.874, we can conclude that 
the Zero-Truncated Negative Binomial Regression model fits better than 
the Negative Binomial Regression model.

Finally, interpret the fitting result.  
The formula of the Zero-Truncated Negative Binomial Regression model is:
$$ Y_{i} \sim NB(\mu_{i}, k) $$
$$ E(Y_{i}) = \mu_{i}, Var(Y_{i}) = \mu_{i}+\frac{\mu_{i}^2}{k}$$
$$ ln(\mu_{i}) = \beta_{0} + \beta_{1}X_{i1} +...+ \beta_{p}X_{ip}$$
From the summary of the Zero-Truncated Negative Binomial Regression model,
we have:  
The value of the coefficient for *SexI*, -0.14064 suggests that 
the log count of Rings for Infant Abalone is 0.14064 less than Female Abalone.  
The value of the coefficient for *SexM*, -0.02083 suggests that 
the log count of Rings for Male Abalone is 0.02083 less than Female Abalone.  
The value of the coefficient for *Length*, 1.50445 suggests that 
the log count of Rings increases by 1.50445 for each mm increase 
in Longest shell measurement.  
The value of the first intercept 1.63153 is the log count of the Rings 
when all predictors equal zero.  
The value of the second intercept 5.19065 is the value of 
the over dispersion parameter *k*.

## SAS


## Stata


## References
1. http://highstat.com/index.php/mixed-effects-models-and-extensions-in-ecology-with-r
2. Wikipedia: Truncated regression model